Ricci-Ollivier Curvature of two Random Walks on the Rooted Phylogenetic Subtree-Prune-Regraft Graph

### 2015-03-30 curvature.tex
a few edits in response to comments

\todo{Chris, especially given that we are at a place that is not known for its CS, I might suggest putting your results concerning curvature first to emphasize to reviewers that this work is not trivial. That's what they will be scanning for after you convince them that the area is interesting.}

@@

Instead\todo{does this seem like a strange comparison to you?}, we consider the distribution of rSPR moves in terms of the number of nodes contained within a subtree.
Recall that a tree with $n$ leaves has $2n-1$ total nodes (ignoring the artificial $\rho$ node).
Given a subtree $R$ with $x$ nodes, observe that there are $2n-1 - x$ possible locations to regraft $R$.
However, some of these moves will result in the same neighboring tree as other rSPR moves.
In particular, where we call the edge connecting the subtree rooted at that node to the rest of the tree the ``node's edge'', we have:
\begin{enumerate}
\item Moving $R$ to its sibling edge results in the same tree, not a neighboring tree,
\item Moving $R$ to its parent edge results in the same tree,
\item Moving $R$ to its grandparent edge is the same as moving its aunt to its sibling edge,
\item Moving $R$ to its aunt edge is the same as moving its aunt to $R$'s edge.
\end{enumerate}
\todo{Is it clear that this is an exhaustive list?}
We can thus assign $2n-1-x-4$ moves to any node at a depth greater than 1 (from the original non-$\rho$ root) and $2n-1-x-2$ moves to depth 1 nodes (lacking both an aunt and grandparent).
Call the portion of the neighborhood of a tree $T$ that can be obtained by moving a subtree $R$ rooted at a node $u$, and that is assigned to $u$, $N(T,u)$.
In this manner, we achieve an alternative solution for computing the degree of a tree in the rSPR graph:

@@

The analysis generating the can be reproduced by simply invoking the SCons (\url{http://scons.org/}) build tool and evaluating cells in an IPython notebook; instructions are in the repository README file.

>>>>

@@

We consider the distribution of rSPR moves in terms of the number of nodes contained within a subtree.
Recall that a tree with $n$ leaves has $2n-1$ total nodes (ignoring the artificial $\rho$ node).
Given a subtree $R$ with $x$ nodes, observe that there are $2n-1 - x$ possible locations to regraft $R$.
However, some of these moves will result in the same neighboring tree as other rSPR moves.
In particular, where we call the edge connecting the subtree rooted at that node to the rest of the tree the ``node's edge'', we have:
\begin{enumerate}
\item Moving $R$ to its sibling edge results in the same tree, not a neighboring tree,
\item Moving $R$ to its parent edge results in the same tree,
\item Moving $R$ to its grandparent edge is the same as moving its aunt to its sibling edge,
\item Moving $R$ to its aunt edge is the same as moving its aunt to $R$'s edge.
\end{enumerate}
We prove in Lemma~\ref{lem:compute_degree} that this list is exhaustive.
We can thus assign $2n-1-x-4$ moves to any node at a depth greater than 1 (from the original non-$\rho$ root) and $2n-1-x-2$ moves to depth 1 nodes (lacking both an aunt and grandparent).
Call the portion of the neighborhood of a tree $T$ that can be obtained by moving a subtree $R$ rooted at a node $u$, and that is assigned to $u$, $N(T,u)$.
In this manner, we achieve an alternative solution for computing the degree of a tree in the rSPR graph:

@@

The analysis can be reproduced by invoking the SCons (\url{http://scons.org/}) build tool and evaluating cells in an IPython notebook; instructions are in the repository README file.
